---
title: "Lab 1 - Data visualization"
format: html
author: "Fardeen Chowdhuru"
editor: visual
---

# Questions

## Part 1

```{r}


library(tidyverse)
install.packages("ggthemes")
library(ggthemes)
midwest

```

**Question 1**

Binwidth = 100

```{r}

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100, fill= "black") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100"
  )
        
```

Binwidth = 1000
```{r}

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 1000, fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 1000"
  )
        
```
```{r}


ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000,fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 10,000"
  )
        
```
```{r}


ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100000,fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100,000"
  )
        
```
I would say Binwidth = 1000 is a great option as It included most of the data in a one side and in a ogranised pattern as if you want to check certain number of pupulation densitity and how many are there in midwest 1000 would be easier than 10,000 or 100 whereas 100000 does not show any represenation of data based on the quanity variable. It shows the shape of the diagram and the skewness of the data. it aslo shows poetential outliers so showing better describing the analysis.

### Question 2
```{r}

ggplot(midwest, aes(x = popdensity)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Population Density of Midwest Counties",
    y = "",
    x = "Population Density (persons per unit area)
    "
  ) +
  theme_minimal() #got it from google for better look for scatterplot diagram

``` 
The data analysis from question 1 and 2 shows that the pop densisty of counties of midwest area is mostly around 0 to 15,000 with median below 6,000. However, it shows there are possible outlier of counties showing some counties with better reesources leading to higher population densisty and as population densisty depend on lot of facots among counties there are many outlies but the data is mostly pushed toward left. Hence, from the data I feel median does not provide much information with having so many outliers and mean would be benefical.

Cook county is one of the outlier with highest pupulation density and far from the median.

### Question 3


```{r}



ggplot(midwest, aes(x = percollege, y = percbelowpoverty))+
  geom_point(aes(color = state, shape= state ))+
  geom_smooth()+
  labs(
    title = 'College Education and Poverty Across Midwest Counties (2000)',
    subtitle = "County-level comparison using U.S. Census data, colored and shaped by state",
    x= "Percent of Adults with a College Degree",
    y= "Percent of Population Below the Poverty Line",
    color = "State",
    shape= "State"
  )+
  scale_x_continuous(labels = scales::label_percent(scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(scale = 1))+  #scale x make its look better as it includes percentage sign. 

  scale_color_colorblind()+
  theme_minimal() #once again it makes it look better
  
 

  
  
```
#According it to the scatter plot - Percent of Adulls with a college degree and population below poverty has somewhat a negative relationship in midwestern states. It indicates that with a rise in college degree among adult the population below poverty line falls. It shows a reationship not a causation. But looking into best with line, the co-relation is strong hence a full conclusion cannot be made. 

#One of the outlier is JACKSON county. with high pupulation of college degree but still a comparative high poverty level. 

### Question 4
```{r}



ggplot(midwest, aes(x = percollege, y = percbelowpoverty))+
  geom_point(aes(color = state, shape= state ))+
  geom_smooth( se = FALSE)+ #I did not use linear method because i feel it should be exponential explaining the best fit line better and i feel it does not work for linear level and showing somewhat relationship with the linear
  labs(
    title = 'College Education and Poverty Across Midwest Counties (2000)',
    subtitle = "County-level comparison using U.S. Census data, colored and shaped by state",
    x= "Percent of Adults with a College Degree",
    y= "Percent of Population Below the Poverty Line",
    color = "State",
    shape= "State"
  )+
  scale_x_continuous(labels = scales::label_percent(scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(scale = 1))+  #scale x make its look better as it includes percentage sign. +
  facet_wrap(~state)+

  scale_color_colorblind()+
  theme_minimal() #once again it makes it look better
  
 

  
  
```
What I prefer among question 3 and question depends on what i am trying to get from the data. IF i want to compare solely among state than I  would choose question 4 but if i want to see the relationship in general i would prefer question 3. However, the graph in question 4 looks more organised to me.
### Question 5
```{r}
ggplot(midwest, aes(x= state, y= area))+
  geom_boxplot()+
  labs(
    title = "Distribution of County Area Sizes Across Midwest States",
    subtitle = "Side-by-side boxplots showing variability in county land area by state",
    x = "State",
    y= "County Area"
  )+
  scale_x_discrete(
    labels = c(
      IL = "Illinois",
      IN = "Indiana",
      MI = "Michigan",
      OH = "Ohio",
      WI = "Wisconsin"
    )
  )+
  
  #i wanted to put full name of states in this plot to make the graph look easier and a small kid can understand it hence I asked chatgpt how can I put state names for x axis and it give me the coe for scale_x_discrete. It helped e me know more about the fuction as well
  
  scale_y_continuous(
    limit = c(0, 0.12),
   
  )+
  #I use this fucntion taking help from chat gpt to show the lable for the higes areaof country on y axis, earlier it was not showing. But now it is showing making the graph analysis more easier for anyone.
  theme_minimal()

```
Michigan and wisconsisn has a higer variability in county sizes with higher interquartlie range and bigger whisker with several high area outliers in indicaitng a spread between high and low county area/

For Indiana and Ohia, it shows the size of country more consistent with more compact data and small interquartile range 

For Illinoise I feel it has moderate variability with bigger interquartile range then Indiana and Ohi. but smaller than michigan and wisconsin. Also compared to michigan and wisconsisn their outlier is withinin the largest county showning less extreme variability.

Michigan has the single largest county. Name of the country is MARQUETTE.

### Question 6
```{r}
midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No"))

ggplot(midwest, aes(x = state,fill = metro))+
  geom_bar(position = "fill")+
  scale_y_continuous(labels = scales::label_percent(scale=1))+
  scale_x_discrete(
    labels = c(
      IL = "Illinois",
      IN = "Indiana",
      MI = "Michigan",
      OH = "Ohio",
      WI = "Wisconsin"
    )
  )+
  labs(
    title = "Proportion of Counties in Metropolitan Areas by State",
    subtitle = "Segmented bar chart showing metro vs non-metro counties",
    x = "State",
    y = "Proportion of Counties",
    fill = "Counties Metropolitan Area"
  )+
  
  theme_minimal()


```
Illinois has slightly more than 25% in metropolital area which is similar with Michigan and wisconsin. 
However, Indiana has slight her than 37.5% counties in metropolitan area. Ohio has the higes counties in metorpolitan area according to the data with more than 40% but below 50%. 
### Question 7
```{r}

ggplot(midwest, aes(x = percollege, y = percbelowpoverty, colour = percwhite))+
  geom_point(size= 2, alpha = 0.5)+
  facet_wrap(~state)+
   labs(
    title = "College Education and Poverty Across Midwest Counties",
    subtitle = "Faceted by state;\nColor indicates percent white population",
    x = "% college educated",
    y = "% below poverty line",
    color = "% white"
  )+
  theme_minimal()
 


```

MENOMINEE county in Wisconsin state is a outlier with way higher bewlo poverty rate of 48.6911% compared to other counties. In this county 10.7% of thew population are white, but the most of the people in the county are from American Indians race with 89%. Whereas the country does not have any black population but there are 5% of other races. This chose the county majority population is american indians whereas whites and other races are minority. 

## Part 2

**Enough about the Midwest!**

### Question 8


### Question 9
