---
title: "Lab 1 - Data visualization"
format: html
author: "Fardeen Chowdhuru"
editor: visual
---

# Questions

## Part 1

```{r}


library(tidyverse)
midwest

```

**Question 1**

Binwidth = 100

```{r}

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100, fill= "black") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100"
  )
        
```

Binwidth = 1000
```{r}

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 1000, fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 1000"
  )
        
```
```{r}


ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000,fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 10,000"
  )
        
```
```{r}


ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100000,fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100,000"
  )
        
```
I would say Binwidth = 1000 is a great option as It included most of the data in a one side and in a ogranised pattern as if you want to check certain number of pupulation densitity and how many are there in midwest 1000 would be easier than 10,000 or 100 whereas 100000 does not show any represenation of data based on the quanity variable. It shows the shape of the diagram and the skewness of the data. it aslo shows poetential outliers so showing better describing the analysis.

### Question 2
```{r}
midwest
ggplot(midwest, aes(x = popdensity)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Population Density of Midwest Counties",
    y = "",
    x = "Population Density (persons per unit area)
    "
  ) +
  theme_minimal() #got it from google for better look for scatterplot diagram

``` 
The data analysis from question 1 and 2 shows that the pop densisty of counties of midwest area is mostly around 0 to 15,000 with median below 6,000. However, it shows there are possible outlier of counties showing some counties with better reesources leading to higher population densisty and as population densisty depend on lot of facots among counties there are many outlies but the data is mostly pushed toward left. Hence, from the data I feel median does not provide much information with having so many outliers and mean would be benefical.

Cook county is one of the outlier with highest pupulation density and far from the median.

### Question 3

### Question 4

### Question 5

### Question 6

### Question 7

## Part 2

**Enough about the Midwest!**

### Question 8

### Question 9
