---
title: "Lab 1 - Data visualization"
format: html
author: "Fardeen Chowdhuruy"
student_id: 1773707

editor: visual
---

# Questions

## Part 1
Data Setup
```{r}


library(tidyverse)
install.packages("ggthemes")
library(ggthemes)
library(scales)
midwest

```

**Question 1**

Binwidth = 100

```{r}

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100, fill= "black") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100"
  )
        
```

Binwidth = 1000
```{r}

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 1000, fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 1000"
  )
        
```
```{r}


ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000,fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 10,000"
  )
        
```
```{r}


ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100000,fill= "black", colour= "white") +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100,000"
  )
        
```
I would say Binwidth = 1000 is a great option as It included most of the data in a one side and in a ogranised pattern as if you want to check certain number of pupulation densitity and how many are there in midwest 1000 would be easier than 10,000 or 100 whereas 100000 does not show any represenation of data based on the quanity variable. It shows the shape of the diagram and the skewness of the data. it aslo shows poetential outliers so showing better describing the analysis.

### Question 2
```{r}
<<<<<<< HEAD

=======
midwest
>>>>>>> be2c1f1f04890e28c606174594edc9d5d311f2b6
ggplot(midwest, aes(x = popdensity)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Population Density of Midwest Counties",
    y = "",
    x = "Population Density (persons per unit area)
    "
  ) +
  theme_minimal() #got it from google for better look for scatterplot diagram
<<<<<<< HEAD

``` 
The data analysis from question 1 and 2 shows that the pop densisty of counties of midwest area is mostly around 0 to 15,000 with median below 6,000. However, it shows there are possible outlier of counties showing some counties with better reesources leading to higher population densisty and as population densisty depend on lot of facots among counties there are many outlies but the data is mostly pushed toward left. Hence, from the data I feel median does not provide much information with having so many outliers and mean would be benefical.

Overall, the population densisty of midwest counties is has low variability withmost data pushed on left with may otliers and showing the vaiarbiliry with the average spread. 

Cook county is one of the outlier with highest pupulation density and far from the median.
=======
>>>>>>> be2c1f1f04890e28c606174594edc9d5d311f2b6

``` 
The data analysis from question 1 and 2 shows that the pop densisty of counties of midwest area is mostly around 0 to 15,000 with median below 6,000. However, it shows there are possible outlier of counties showing some counties with better reesources leading to higher population densisty and as population densisty depend on lot of facots among counties there are many outlies but the data is mostly pushed toward left. Hence, from the data I feel median does not provide much information with having so many outliers and mean would be benefical.
### Question 3


```{r}



ggplot(midwest, aes(x = percollege, y = percbelowpoverty))+
  geom_point(aes(color = state, shape= state ))+
  geom_smooth()+
  labs(
    title = 'College Education and Poverty Across Midwest Counties (2000)',
    subtitle = "County-level comparison using U.S. Census data, colored and shaped by state",
    x= "Percent of Adults with a College Degree",
    y= "Percent of Population Below the Poverty Line",
    color = "State",
    shape= "State"
  )+
  scale_x_continuous(labels = scales::label_percent(scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(scale = 1))+  #scale x make its look better as it includes percentage sign. 

  scale_color_colorblind()+
  theme_minimal() #once again it makes it look better
  
 

  
  
```
#According it to the scatter plot - Percent of Adulls with a college degree and population below poverty has somewhat a negative relationship in midwestern states. It indicates that with a rise in college degree among adult the population below poverty line falls. It shows a reationship not a causation. But looking into best with line, the co-relation is strong hence a full conclusion cannot be made. 

#One of the outlier is JACKSON county. with high pupulation of college degree but still a comparative high poverty level. 

### Question 4
```{r}



ggplot(midwest, aes(x = percollege, y = percbelowpoverty))+
  geom_point(aes(color = state, shape= state ))+
  geom_smooth( se = FALSE)+ #I did not use linear method because i feel it should be exponential explaining the best fit line better and i feel it does not work for linear level and showing somewhat relationship with the linear
  labs(
    title = 'College Education and Poverty Across Midwest Counties (2000)',
    subtitle = "County-level comparison using U.S. Census data, colored and shaped by state",
    x= "Percent of Adults with a College Degree",
    y= "Percent of Population Below the Poverty Line",
    color = "State",
    shape= "State"
  )+
  scale_x_continuous(labels = scales::label_percent(scale = 1)) +
  scale_y_continuous(labels = scales::label_percent(scale = 1))+  #scale x make its look better as it includes percentage sign. +
  facet_wrap(~state)+

  scale_color_colorblind()+
  theme_minimal() #once again it makes it look better
  
 

  
  
```
What I prefer among question 3 and question depends on what i am trying to get from the data. IF i want to compare solely among state than I  would choose question 4 but if i want to see the relationship in general i would prefer question 3. However, the graph in question 4 looks more organised to me.
### Question 5
```{r}
ggplot(midwest, aes(x= state, y= area))+
  geom_boxplot()+
  labs(
    title = "Distribution of County Area Sizes Across Midwest States",
    subtitle = "Side-by-side boxplots showing variability in county land area by state",
    x = "State",
    y= "County Area"
  )+
  scale_x_discrete(
    labels = c(
      IL = "Illinois",
      IN = "Indiana",
      MI = "Michigan",
      OH = "Ohio",
      WI = "Wisconsin"
    )
  )+
  
  #i wanted to put full name of states in this plot to make the graph look easier and a small kid can understand it hence I asked chatgpt how can I put state names for x axis and it give me the coe for scale_x_discrete. It helped e me know more about the fuction as well
  
  scale_y_continuous(
    limit = c(0, 0.12),
   
  )+
  #I use this fucntion taking help from chat gpt to show the lable for the higes areaof country on y axis, earlier it was not showing. But now it is showing making the graph analysis more easier for anyone.
  theme_minimal()

```
Michigan and wisconsisn has a higer variability in county sizes with higher interquartlie range and bigger whisker with several high area outliers in indicaitng a spread between high and low county area/

For Indiana and Ohia, it shows the size of country more consistent with more compact data and small interquartile range 

For Illinoise I feel it has moderate variability with bigger interquartile range then Indiana and Ohi. but smaller than michigan and wisconsin. Also compared to michigan and wisconsisn their outlier is withinin the largest county showning less extreme variability.

Michigan has the single largest county. Name of the country is MARQUETTE.

### Question 6
```{r}
midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No"))

ggplot(midwest, aes(x = state,fill = metro))+
  geom_bar(position = "fill")+
  scale_y_continuous(labels = scales::label_percent(scale=1))+
  scale_x_discrete(
    labels = c(
      IL = "Illinois",
      IN = "Indiana",
      MI = "Michigan",
      OH = "Ohio",
      WI = "Wisconsin"
    )
  )+
  labs(
    title = "Proportion of Counties in Metropolitan Areas by State",
    subtitle = "Segmented bar chart showing metro vs non-metro counties",
    x = "State",
    y = "Proportion of Counties",
    fill = "Counties Metropolitan Area"
  )+
  
  theme_minimal()


```
Illinois has slightly more than 25% in metropolital area which is similar with Michigan and wisconsin. 
However, Indiana has slight her than 37.5% counties in metropolitan area. Ohio has the higes counties in metorpolitan area according to the data with more than 40% but below 50%. 
### Question 7
```{r}

ggplot(midwest, aes(x = percollege, y = percbelowpoverty, colour = percwhite))+
  geom_point(size= 2, alpha = 0.5)+
  facet_wrap(~state)+
   labs(
    title = "College Education and Poverty Across Midwest Counties",
    subtitle = "Faceted by state;\nColor indicates percent white population",
    x = "% college educated",
    y = "% below poverty line",
    color = "% white"
  )+
  theme_minimal()
 


```

MENOMINEE county in Wisconsin state is a outlier with way higher bewlo poverty rate of 48.6911% compared to other counties. In this county 10.7% of thew population are white, but the most of the people in the county are from American Indians race with 89%. Whereas the country does not have any black population but there are 5% of other races. This chose the county majority population is american indians whereas whites and other races are minority. 

## Part 2

**Enough about the Midwest!**

```{r}
nc_county <- read_csv("data/nc-county.csv")

```
<!-- county: Name of county. -->
<!-- state_abb: State abbreviation (NC). -->
<!-- state_name: State name (North Carolina). -->
<!-- land_area_m2: Land area of county in square metres, based on the 2020 census. -->
<!-- land_area_mi2: Land area of county in square miles, based on the 2020 census. -->
<!-- population: Population of county, based on the 2020 census. -->
<!-- density: Population density calculated as population divided by land area in square miles. -->
### Question 8
Before doing the coding i have guessed that popution density and land area has a negative relationship as population densitty should increase when land area decreases.

```{r}

ggplot(nc_county, aes(x = land_area_mi2, y= density))+
  geom_point(colour= "darkblue")+
  geom_smooth(se = FALSE, color = "darkred")+
  labs(
    title = "Relationship Between County Land Area (Square Miles) and Population Density",
    subtitle = "Scatter plot of counties in North Carolina (2020 Census)",
    x = "Land Area\n(square miles)",
    y = "Population Density\n(people per square mile)"
  )+
  scale_x_continuous(limits = c(150,1000),
  breaks = seq(100, 1000, by = 100))+ # i wanted my grid to be by 100 till 950 so that it is easy for viewrs to see the data and compare for this one I asked chatgpt to how to do that and it give my this conde and it is pretty understandable

  scale_y_continuous(limits = c(0,2500))+
  
  theme_light() #i like the plot design using theme_light instead of theme_minimal
  
```
After analysing the data and making a scatter plot, I would so if you just see the best fit line, my guess of a negative relationship was not correct. This is becase even though land area is increasing, most of the county's population density is not in decreasing strongly. insteade the population density remains at the similar level despite county with higher land area. Hence, it can be said that population of a country does not strongly depend on Land Area. 

### Question 9
```{r}

ggplot(nc_county, aes(x = land_area_m2, y= density))+
  geom_point(colour= "darkgreen")+
  geom_smooth(se = FALSE, color = "darkred")+
  labs(
    title = "Relationship Between County Land Area(square metres) and Population Density",
    subtitle = "Scatter plot of counties in North Carolina (2020 Census)",
    x = "Land Area\n(square metres)",
    y = "Population Density\n(people per square mile)"
  )+
  scale_x_continuous(labels = label_number(scale = 1e-6, suffix = " Million")) +
   scale_y_continuous(limits = c(0,2500))+

  theme_light() #i like the plot design using theme_light instead of theme_minimal
  

```
Relationship Status is same meaning populuation density (people per square mile)- is has no strong relation with Land Area of square metres. I feel the reason it is showing similar is because in the population density because the density does not only depend on the land area. The number of people living the county( population) matter which did not change in this analysis.
